Optimal design of experiment

### 2024-01-22 SFE_article_Design_of_Experiment/Sections/Literature_Review.tex
22.01

	This study aims to analyze the influence of changes in operating conditions on the supercritical extraction process described in {\color{red}article 1}. The emphasis is put on the effect of the mass flow rate, pressure, and the inlet temperature. The relation between input and output is obtained by applying a sensitivity analysis. Sensitivity analysis examines the impact of varying inputs or model parameters on the system's output. The aim is to understand and to allocate the source of uncertainty in the output to the corresponding inputs or parameters. There are many sensitivity analysis methods, which include but are not limited to those listed below:

	\begin{itemize}
		\item One-at-a-time method
		\item Derivative-based local methods
		\item Variance-based methods
	\end{itemize}

	Different supercritical extraction models were analyzed using sensitivity analysis. \citet{Fiori_2007}, performed the sensitivity calculations by varying the parameters within their confidence interval and observing how the model results changed. This allows to evaluate the effect of the uncertainties on model predictions. The sensitivity analysis revealed that the particle diameter and the internal mass transfer coefficient are significant for the extraction process. The effect of changing some operative conditions was also investigated, underlining how the solvent flow rate and the seed milling affect the extraction process.

	\citet{Santos2000}, in their work, considered a model of supercritical extraction process for semi-continuous isothermal and isobaric extraction process using carbon dioxide as a solvent. The parametric sensitivity analysis was carried out by applying disturbances of 10\% in the values of the normal operation conditions.

	\citet{Hatami2024} used a one-factor-at-a-time sensitivity analysis to assess the response of NPV concerning variations in both technical and economic variables. Their findings show that the most influential factors on NPV include the price of the extract, the interest rate, the dynamic time of SFE, and the project lifetime.

	\citet{Poletto1996} provided a general dimensionless model for the supercritical extraction process of vegetable and essential oils and applied a sensitivity analysis. They found that a dimensionless partition coefficient and a dimensionless characteristic time appeared as the most important parameters of the extraction process. The sensitivity calculations were performed by varying the parameters and analyzing the model response.

>>>>

	The design of experiments (DOE) also known as experiment design or experimental design, is the design of any task that aims to describe and explain the variation of information under conditions that are hypothesized to reflect the variation. The term is generally associated with experiments in which the design introduces conditions that directly affect the variation, but may also refer to the design of quasi-experiments, in which natural conditions that influence the variation are selected for observation.

	\citet{Oinas1992} presents how the experimental design methods can be applied for analyzing mass transfer and reaction in stirred gas—liquid and gas—liquid—solid reactors. In this work a sequential approach to experimentation is suggested. First preliminary set of 8 2-level experiments ($2^8$ factorial design) were performed. The identifiability of the parameters was then considered and some reparametrisations done. The precision of the parameter values were further improved with D-optimal experiments.

	\citet{Kassama2008} used the response surface methodology with the central composite rotatetable design (CCRD) model. The goal of this work was to optimize parameters for supercritical carbon dioxide extraction of lycopene ($C_{40}H_{56}$) from dried tomato skin. The CCRD consisting of three-factored factorial design with two levels was used in this study. The factors used were temperature of the extraction chamber (40 and 70 $^\circ C$), pressure of the extraction fluid (25 and 45 MPa), and modifier concentration (5 and 15\%). A second-degree polynomial equation was developed from a response surface analysis for all trans-lycopene yield and the highest yield was predicted at 62 $^\circ C$, 45 MPa (450 bar) and 14\% temperature, pressure and modifier concentration, respectively and the recovery of all trans-lycopene was 33\%.

	\citet{Abrahamsson2016} selected extraction conditions based on a symmetrical full factorial design of experiments with three factors and three levels. The factors were pressure ranging from 15 to 30 MPa, temperature from 313 to 353 K and compressed $CO_2$ pump volumetric flow rate from 0.5 to 1.5 mL/min, with three replicates at the center point (22.5 MPa, 333 K, 1 mL/min) thus resulting in a total of 29 extractions. By including all of the 29 experiments for the model calibration, a more robust estimates of parameters were obtained.

### 2024-01-22 SFE_article_Design_of_Experiment/Sections/Optimal_design_of_experiment.tex
22.01

	The design of experiments (DOE) also known as experiment design or experimental design, is the design of any task that aims to describe and explain the variation of information under conditions that are hypothesized to reflect the variation. The term is generally associated with experiments in which the design introduces conditions that directly affect the variation, but may also refer to the design of quasi-experiments, in which natural conditions that influence the variation are selected for observation.

	\subsubsection{Maximum likelihood}
	%Let $Y$ be the vector of all experimental results (random vector) to be used to estimate parameters $p$ (where $p$ is the subset of parameter space $\Theta$) and model output $y(p)$. The vector of the corresponding quantities computed by the model $\dot{x}(t)=G(x(t);p)$. For a parallel model, the parameters of which are to be estimated from measurements of an output vector $\dot{y}=g(x)$. We shall call any optimizer of $j$, i.e., any $\hat{p}$ that corresponds to an optimal value of the cost function $j$, an estimate of $p$ in the sense of $j$.

	The vector $\hat{p}_{ml}$ will be a maximum-likelihood estimate if it maximizes the cost function:

@@

	If $p$ were fixed, $\pi_y(Y|p)$ would be the probability density of the random vector $Y$ generated by a model with parameters $p$. Here, to the contrary, $Y$ is fixed and corresponds to the observations. Considered as a function of $p$, $\pi_y(Y|p)$ is then called the likelihood of $Y$. The maximum-likelihood method looks for the parameter vector $p$ value that gives the observed data the highest likelihood. In practice, it is often easier to look for $\hat{p}_{ml}$ by maximizing the log-likelihood function, yielding the same estimate since the logarithm function is monotonically increasing.

@@

	where the vector $y(t_i, p^*)$ is the output of a deterministic model, $p^*$ is the true value of the parameter vector, and $\epsilon_i$ belongs to a sequence of independent random variables with probability density $\pi_{\epsilon}(\epsilon_i)$. Since the $\epsilon_i$ are independent

@@

	{\footnotesize
	\begin{equation}
		e^y(t_i, p) = Y(t_i) - y(t_i, p)
	\end{equation} }

	For the true value of the parameters, it satisfies $e^y(t_i, p^*)=\epsilon_i$. and, since $y$ is deterministic, $\pi_{y_i} (Y(t_i)|p) = \pi_{\epsilon_i} (e^Y(t_i)|p)$. The likelihood of the $n_t$ observations can be written as

@@

	Considering the mismatch between the dataset and the model output, the expectation $\mathop{\mathbb{E}}_{Y|p}$ can be redefined as $\mathop{\mathbb{E}}_{Y|p} \left[ \left( Y(t_i) - y(t_i, p) \right) \left( Y(t_k) - y(t_k, p) \right) \right] = \mathop{\mathbb{E}}_{Y|p} \left[ \epsilon_i \epsilon_k \right]$. For $i=k$, the expectation of $\epsilon_i$ is the variance of the noise $\sigma_{t_i}^2$. Analogously, the expectation becomes 0 if $i \neq k$ because the measurement noises are independent at different times.

@@

	Let $\hat{p}$ be an (absolutely) unbiased estimator $p^*$, i.e. such that it were possible to replicate the same experiment and estimate $\hat{p}$ an infinite number of times, the mean of the estimates would coincide with the true value. Let $P$ be the covariance matrix of this estimator. Since $\hat{p}$ is unbiased, $P$ can be written as

@@

	In other words, the precision to which we can estimate $p$ is fundamentally limited by the Fisher information of the likelihood function. Based on the Cramer-Rao inequality, the Fisher information matrix can used to calculate the covariance matrices associated with maximum-likelihood estimates.

@@

	Let's consider that each scalar observation in our study can be expressed as $y(\xi_i)$, where the $n_\xi$-dimensional vector $\xi_i$ representing the specific experimental conditions (such as the timing of measurements, operating conditions, etc.) under which the i'th observation is gathered. When collecting $n_t$ such observations, the assembly of these $\xi_i$ vectors forms the matrix $\Xi = (\xi_1, \xi_2,..., \xi_{n_t})$, which combine all the experimental conditions that need optimization. In order to align the design of the experiment with practical realities, it's important to take into account various constraints, such as the total duration of the experiments, the maximum temperature of the inlet stream, and the minimum interval between sampling events. The set of all possible combinations for $\bar{\Xi}$ that adhere to these constraints is denoted as $\Xi$.

@@

	The cost function should describe the amount of information from an experiment. For that purpose it can be assumed a function $\phi$ can be related to the Fisher information obtained an arbitrary operating conditions.

	{\footnotesize
	\begin{equation}
		j(\Xi) = \phi\left[ \text{FIM(p)}(p, \Xi) \right]
	\end{equation} }

@@

	{\footnotesize
	\begin{align}
		\phi_k(\text{FIM(p)}) &= \left[\frac{1}{n_p} \text{trace}\left( Q\text{FIM(p)}^{-1}Q^\top \right)^k \right]^{1/k} \quad &\text{if} \det F \neq 0 \nonumber \\
		\phi_k(\text{FIM(p)}) &= \infty \quad &\text{if} \det F = 0
	\end{align} }

>>>>

	\subsubsection{Maximum likelihood}

	Following the work of \citet{Walter2010}, the vector $\hat{p}_{ml}$ will be a maximum-likelihood estimate if it maximizes the cost function:

@@

	If $p$ were fixed, $\pi_y(Y|p)$ would be the probability density of the random vector $Y$ generated by a model with parameters $p$ and observations $Y$. Considered as a function of $p$, $\pi_y(Y|p)$ is then called the likelihood of $Y$. The maximum-likelihood method looks for the parameter vector $p$ value that gives the observed data the highest likelihood. In practice, it is often easier to look for $\hat{p}_{ml}$ by maximizing the log-likelihood function, yielding the same estimate since the logarithm function is monotonically increasing.

@@

	where the vector $y(t_i, p^*)$ is the output of a deterministic model, $p^*$ is the true value of the parameter vector, and the error $\epsilon_i$ belongs to a sequence of independent random variables with probability density $\pi_{\epsilon}(\epsilon_i)$. Since the $\epsilon_i$ are independent

@@

	{\footnotesize
	\begin{equation} \label{EQ:Error_measurment}
		e^y(t_i, p) = Y(t_i) - y(t_i, p)
	\end{equation} }

	For the true value of the parameters, it satisfies $e^y(t_i, p^*)=\epsilon_i$. and, since $y$ is deterministic, $\pi_{y_i} (Y(t_i)|p) = \pi_{\epsilon_i} (e^y(t_i),p)$. The likelihood of the $n_t$ observations can be written as

@@

	Considering the mismatch between the dataset and the model output, the Equation \ref{EQ:Error_measurment} can be used too redefine the expectation $\mathop{\mathbb{E}}_{Y|p}$ as $\mathop{\mathbb{E}}_{Y|p} \left[ \left( Y(t_i) - y(t_i, p) \right) \left( Y(t_k) - y(t_k, p) \right) \right] = \mathop{\mathbb{E}}_{Y|p} \left[ \epsilon_i \epsilon_k \right]$. For $i=k$, the expectation of $\epsilon_i$ is the variance of the noise $\sigma_{t_i}^2$. Analogously, the expectation becomes 0 if $i \neq k$ because the measurement noises are independent at different times.

@@

	Let $\hat{p}$ be an (absolutely) unbiased estimator $p^*$, i.e. such that it were possible to replicate the same experiment and estimate $\hat{p}$ an infinite number of times, the mean of the estimates would coincide with the true value. Let $\Sigma$ be the covariance matrix of this estimator. Since $\hat{p}$ is unbiased, $\Sigma$ can be written as

@@

	In other words, the precision to which we can estimate $p$ is limited by the Fisher information of the likelihood function. Based on the Cramer-Rao inequality, the Fisher information matrix can used to calculate the covariance matrices associated with maximum-likelihood estimates.

@@

	Let's consider that each scalar observation in a study can be expressed as $y(\xi_i)$, where the $n_\xi$-dimensional vector $\xi_i$ representing the specific experimental conditions (such as the timing of measurements, operating conditions, etc.) under which the i'th observation is gathered. When collecting $n_t$ such observations, the assembly of these $\xi_i$ vectors forms the matrix $\Xi = (\xi_1, \xi_2,..., \xi_{n_t})$, which combine all the experimental conditions that need optimization. In order to align the design of the experiment with practical realities, it's important to take into account various constraints, such as the total duration of the experiments, the maximum temperature of the inlet stream, and the minimum interval between sampling events. The set of all possible combinations for $\bar{\Xi}$ that adhere to these constraints is denoted as $\Xi$.

@@

	The cost function should describe the amount of information from an experiment. For that purpose it can be assumed a function $\kappa$ can be related to the Fisher information obtained an arbitrary operating conditions.

	{\footnotesize
	\begin{equation}
		j(\Xi) = \kappa\left[ \text{FIM}(p, \Xi) \right]
	\end{equation} }

@@

	{\footnotesize
	\begin{align}
		\kappa_k(\text{FIM(p)}) &= \left[\frac{1}{n_p} \text{trace}\left( Q\text{FIM}()^{-1}Q^\top \right)^k \right]^{1/k} \quad &\text{if} \det \text{FIM} \neq 0 \nonumber \\
		\kappa_k(\text{FIM(p)}) &= \infty \quad &\text{if} \det \text{FIM} = 0
	\end{align} }
